---
title: "STA 322 Project 1: A Survey of Trinity College Faculty"
author: "NAME 1 and NAME 2"
date: "October 9, 2026"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(dplyr)
library(readr)
library(survey)
library(knitr)
```

# 1. Population and sampling frame

**Target population.** Tenured and tenure-track faculty (Assistant, Associate, and Full Professors)
whose *primary* appointment is in one of the 27 departments of Trinity College of Arts & Sciences,
as listed on Scholars@Duke in October 2026. Excluded: emeriti; professors of the practice;
lecturers and instructors; research, adjunct, and visiting faculty; and people whose primary
appointment is outside Trinity (e.g., School of Medicine faculty with a secondary Trinity appointment).
Interdisciplinary programs (e.g., Linguistics, Dance) contain no eligible primary appointments, so
the population is the same whether or not they are counted.

**Frame construction.** Starting from the Scholars@Duke people search filtered to
*Trinity College of Arts & Sciences* and *Faculty* (1,440 people), we recorded only each person's
profile link, name and headline title. We assigned each person to one department using the primary
appointment recorded in Scholars, and we assigned departments to divisions using Trinity's own
division pages. This left $N = 495$ eligible faculty in 27 departments. No survey variables were
collected for anyone outside the sample.

```{r frame}
frame <- read_csv("frame.csv", show_col_types = FALSE)
kable(frame |> select(division, dept, M_i), col.names = c("Division", "Department", "Eligible faculty"))
```

# 2. Survey design and justification

We used a **stratified simple random sample**, with the 27 departments as strata.

* *Why stratify by department?* The fraction of women (our primary target) plausibly differs a lot
  across fields. Stratifying removes between-department variation from the sampling error, and
  proportional allocation can never do worse than an SRS of the same size.
* *Why not cluster sampling?* Clustering saves cost when reaching a cluster is the expensive step.
  Here every sampled person is a separate profile lookup, and the full frame was already available,
  so sampling only some departments would lose precision without saving any time.
* *Allocation.* Proportional to department size, targeting $n = 120$, with a minimum of 2 per
  department so that every within-stratum variance can be estimated. Rounding and the minimum give
  $n = 124$.

```{r sample}
dat <- read_csv("data.csv", show_col_types = FALSE)
kable(dat |> distinct(dept, M_h, n_h, pi, w) |> arrange(dept),
      digits = 3, col.names = c("Department", "$N_h$", "$n_h$", "$\\pi_i$", "$w_i$"))
```

# 3. Inclusion probabilities and weights

Within department $h$, each eligible person is selected with probability $\pi_i = n_h / N_h$, so
their weight is $w_i = N_h / n_h$. The weights sum to $N = 495$. Because of the minimum of 2 per
stratum, small departments are slightly oversampled, so the weights are not all equal. The
Horvitz–Thompson estimator therefore reduces to $\sum_h (N_h/N)\,\bar y_h$. We use the exact
stratified-SRS variance with a finite population correction in each stratum.

```{r design}
dat <- dat |>
  mutate(female = as.integer(sex == "F"),
         phd_duke = ifelse(has_phd == 1, phd_duke, NA))
stopifnot(sum(dat$w) - 495 < 1e-8)
des <- svydesign(id = ~1, strata = ~dept, fpc = ~M_h, data = dat)
```

# 4. Sample size

We sampled 124 of 495 eligible faculty (25%). Planning on the worst case $p = 0.5$ and ignoring
the stratification gain, the 95% margin of error for Q1 is about $\pm 7.6$ percentage points.

# 5. Data collection and coding rules

Data were taken from each sampled person's Scholars@Duke record, reading the public profile and
its official data feed, in October 2026.

| Variable | Rule |
|---|---|
| `sex` | Coded by us from the profile (photo, name, pronouns in bio). `pronoun_hint` counts pronouns in the bio as an aid. |
| `grant_2026` | 1 if any listed grant, fellowship or gift has a start-to-end interval overlapping calendar year 2026; 0 otherwise, including when none are listed. |
| `has_phd` | 1 if a Ph.D. or D.Phil. is listed; 0 if education is listed without a PhD (e.g., MFA, JD, DMA); missing if no education is listed. |
| `years_since_phd` | 2026 minus the PhD year (the earliest PhD if more than one is listed). |
| `phd_duke` | 1 if the PhD is from Duke University; defined only for PhD holders. |

# 6. Results

```{r results}
ci_prop <- function(f, d) {
  est <- svyciprop(f, d, method = "logit")   # logit CI keeps bounds inside [0, 1]
  c(estimate = unname(coef(est)), SE = unname(SE(est)), lower = confint(est)[1], upper = confint(est)[2])
}
phd_des <- subset(des, has_phd == 1)
q3 <- svymean(~years_since_phd, phd_des, na.rm = TRUE)

res <- rbind(
  "Q1. Fraction women"                       = ci_prop(~female, des),
  "Q2. Fraction with grant ongoing in 2026"  = ci_prop(~grant_2026, des),
  "Q3. Mean years since PhD (PhD holders)"   = c(coef(q3), SE(q3), confint(q3)),
  "Q4. Fraction of PhD holders with Duke PhD" = ci_prop(~phd_duke, phd_des)
)
kable(res, digits = 3, col.names = c("Estimate", "SE", "95% CI lower", "95% CI upper"))
```

For comparison, the unweighted sample results:

```{r naive}
dat |> summarise(female = mean(female), grant_2026 = mean(grant_2026),
                 years_since_phd = mean(years_since_phd[has_phd == 1], na.rm = TRUE),
                 phd_duke = mean(phd_duke, na.rm = TRUE)) |> kable(digits = 3)
```

*Interpretation: TO WRITE.*

Optional breakdown by division, using domain estimates:

```{r by-division}
svyby(~female + grant_2026, ~division, des, svymean) |> kable(digits = 3)
```

# 7. Limitations

* **Measurement error in sex.** Sex was judged by us from public profile information, not
  self-reported, and may be misclassified.
* **Grant undercount.** Scholars lists the grants faculty choose to display, and some profiles show
  none. Unlisted grants are coded 0, so `grant_2026` is a lower bound.
* **Grant dates.** We counted any award whose listed interval overlaps 2026. Records with missing or
  outdated end dates may be misclassified.
* **Frame coverage.** The frame is only as current as Scholars@Duke. People with outdated titles or
  primary appointments, or without a profile, are misclassified or missing.
* **Missing education.** Profiles without education data are excluded from Q3 and Q4, which
  assumes they are missing at random.
* **Small strata.** Several departments have only 2 sampled people, so their within-stratum
  variance estimates are unstable.

# Appendix: code

Frame construction (`build_frame.py`, `annotate_frame.py`, `map_divisions.py`), sample selection
(`sample.R`, seed 21) and data collection (`collect.py`) are included with the submission.

```{r appendix-sample, code=readLines("sample.R"), eval=FALSE}
```
